### 2018-04-06 paper/approach.tex
editing approach, need to rework overview, and move some appendix into main paper

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd{ \mathcal{R} }(\mathcal{S})$ %
    \label{line:solve-begin}}{ %
    % recurse on a partial solution to the CHC system:
    \myproc{$\solveaux( \mathcal{ \sigma } )$ %
      \label{line:aux-begin}}{
      % construct predicates that:
      $\mathcal{R}' \assign \setformer{ R }{ %
        % are not defined by sigma,
        R \notin \domainof{ \sigma }, %
        \bigland_{ (R, R') \in \depsof{ \mathcal{S} } } 
        R' \in \domainof{ \sigma } }$ \label{line:cons-nodeps} \;
      % case split: on D:
      \Switch{$\mathcal{R}'$ \label{line:case-deps}}{ %
        \lCase{$\emptyset$:}{ %
          % D is empty: return sigma as solution
          \Return{$\sigma$} \label{line:ret-soln} } %
        \Case{$\elts{ R } \disjunion \mathcal{R}''$: %
          \label{line:case-frontier}}{ %
          % D contains an element R. Update sigma to map R to result
          % of interpolation query on suitable formulas.
          \Return{$\solveaux( %
            \upd{\sigma}{R}{ %
              \solveitp(\prectr{ \mathcal{R} }( %
              \mathcal{S}, \sigma, R), %
              \postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R))})$} %
          \label{line:ret-recurse} \;
        } %
      } %
    \label{line:solve-aux-end} } %
    % check if query is feasible:
    \lIf{$\issat(\vc{ \mathcal{R} }(\mathcal{S}))$}{ % 
      \Return{$\nosoln$} \label{line:check-sat} }
    % return result of calling aux on base case:
    \Return{$\solveaux(\emptyset)$} \label{line:aux-empty}
    \; }
  %
  \caption{$\solvecdd{ \mathcal{R} }$: for $\mathcal{R} \in \relpreds$
    and CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
    solution to $\mathcal{S}$ or the value $\none$ to denote that
    $\mathcal{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}
%
For $\mathcal{R} \in \relpreds$, the procedure $\solvecdd{ \mathcal{R}
}$ (\autoref{alg:solve-cdd}) is a solver for CDD systems in $\chcs{
  \mathcal{R} }$.
%
$\solvecdd{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R}
}$, defines a procedure $\solveaux$ that, given a partial solution
$\sigma$ of $\mathcal{S}$, returns an extension of $\sigma$ that is a
solution of $\mathcal{S}$
(\autoref{line:aux-begin}---\autoref{line:solve-aux-end}).
%
$\solvecdd{ \mathcal{R} }$ runs a procedure $\vc{ \mathcal{R} }$ that,
given a CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
formula that is satisfiable if and only if $\mathcal{S}$ has no
solution, which we refer to as a \emph{counterexample
  characterization} of $\mathcal{S}$ (\autoref{line:check-sat};
%
the implementation of $\vc{ \mathcal{R} }$ is given below).
%
$\solvecdd{ \mathcal{R} }$ runs $\issat$ (\autoref{sec:itps}) on $\vc{
  \mathcal{R} }(\mathcal{S})$;
%
if $\issat$ returns that $\vc{ \mathcal{R} }(\mathcal{S})$ is
satisfiable, then $\solvecdd{ \mathcal{R} }$ returns that
$\mathcal{S}$ has no solution (\autoref{line:check-sat}).
% how the aux function is used:
Otherwise, $\solvecdd{ \mathcal{R} }$ runs $\solveaux$ on the empty
partial solution and returns the result (\autoref{line:aux-empty}).

% walk through the aux function:
$\solveaux$, given partial solution $\sigma$, constructs the set
$\mathcal{R}'$ of each relational predicate $R$ such that $R$ is not
in the domain of $\sigma$ and all dependents of $R$ in $\mathcal{S}$
are in the domain of $\sigma$ (\autoref{line:cons-nodeps}).
%
If $\mathcal{R}'$ is empty, then $\solveaux$ returns $\sigma$ as a
solution of $\mathcal{S}$ (\autoref{line:ret-soln}).
%
Otherwise, if there is some $R \in \mathcal{R}'$
(\autoref{line:case-frontier}), then $\solveaux$ extends $\sigma$ to
map $R$ to an interpolant of the formulas $\prectr{ \mathcal{R}
}(\mathcal{S}, \sigma, R)$ and $\postctr{ \mathcal{R} }(\mathcal{S},
\sigma, R)$, defined below.
%
The interpolant of the two formulas is synthesized by running the
procedure $\solveitp$ (\autoref{sec:itps}).
%
$\solveaux$ recurses on the extended partial solution and returns the
result (\autoref{line:ret-recurse}).
%
\begin{ex}
  $\solvecdd{ \mcpreds }$, given $\mcchc$, may generate
  interpretations of $\mcpreds$ in any topological ordering of the
  dependency relation of $\mcchc$, depending on the relational
  predicates that it chooses at \autoref{line:case-deps}.
  %
  One such ordering is \cc{L3}, \cc{L4}, \cc{dbl}, \cc{L9}, \cc{main},
  $\mathsf{Err}$.
\end{ex}

\subsubsection{Constructing the Counterexample Characterization of a
  CDD}
%
For $\mathcal{S'} \in \chcs{ \mathcal{R} }$, the formula $\vc{
  \mathcal{R} }(\mathcal{S'})$ is a compact representation of the
counterexample characterization of $\mathcal{S'}$, defined as follows.
% introduce the isused predicates:
For each $R \in \mathcal{R}$, let there be a Boolean predicate
$\used{R}$.
% for each relational predicate, define constraint:
The constraint of $R$, denoted $\ctrof{R} \in \formulas$, constrains
that if $R$ occurs in a derivation of $\mathcal{S'}$, then there is
some $\mathcal{C} \in \mathcal{S'}$ in which $R$ is the head such that
\textbf{(1)} each $R' \in \mathcal{R}$ in an application in the body
of $\mathcal{C}$ is used and \textbf{(2)} the values assigned to each
application of $R'$ are equal to the values assigned to the head
variables of $R'$ in $\mathcal{S'}$.
%
I.e., $\ctrof{R}$ is
\begin{align*}
% for each clause with R as a head:
\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} }, %
    R \in \mathcal{R}, \\
  (\mathcal{B}, (R, \varsof{ \mathcal{S'} }(R))) \in
  \mathcal{S'} } } %
\left( \ctrof{ \mathcal{B} } \land %
\bigland_{ \substack{R' \in \mathcal{R}, Y \in X^{*}, \\
    (R', Y) \in \appsof{ \mathcal{B} } } } %
\used{R'} \land \varsof{ \mathcal{S'} }(R') = Y \right)
\end{align*}
%
The counterexample characterization of $\mathcal{S'}$ constrains that
\textbf{(1)} the query predicate of $\mathcal{S'}$ must be used in a
derivation and %
\textbf{(2)} for $R \in \mathcal{R}$, if $R$ is used, then its
constraint must hold.
%
I.e., $\vc{ \mathcal{R} }(\mathcal{S'})$ is $\used{
  \queryof{\mathcal{S'} } } \land %
\bigland_{R \in \mathcal{R} } \left( \used{R} \implies \ctrof{R}
\right)$.

\begin{ex}
  % 
  $\vc{ \mcpreds }(\mcchc)$ characterizes all runs of \cc{multiCall}
  that result in an error.
  %
  Although the relational predicates \cc{L4} and \cc{dbl} occur in the
  bodies of multiple clauses, they are each modeled with only a single
  instance of logical variables.
  % 
  Such a collection is sufficient because each run of \cc{multiCall}
  reaches \cc{L4} and \cc{dbl} at most once.
  %
  Correspondingly, each derivation of $\mcchc$ contains at most one
  occurrence of relational predicates \cc{L4} and \cc{dbl}.
\end{ex}

% pre-constraint:
\subsubsection{Constructing the pre-constraint of a relational predicate}
\label{sec:cons-pre}
%
For partial solution $\sigma : \mathcal{R} \to \formulas$ of
$\mathcal{S}$, and $R \in \mathcal{R}$ for which each dependency is in
the domain of $\sigma$, the pre-constraint $\prectr{ \mathcal{R}
}(\mathcal{S}, \sigma, R)$ is a formula that is entailed by the
interpretation of $R$ in each solution of $\mathcal{S}$ that extends
$\sigma$.
%
In particular, $\prectr{ \mathcal{R} }(\mathcal{S}, \sigma, R)$ is:
\[
\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} } \\
  (\mathcal{B}, (R, \varsof{ \mathcal{S} }(R) ) ) \in \mathcal{S} } }
\left( \ctrof{ \mathcal{ B } } \land %
\bigland_{ A \in \appsof{ \mathcal{B} } } %
\sigma( \relof{A} )[ \argsof{ A } ] \right)
\]
%
\begin{ex}
  \label{ex:pre-ctr}
  Assume that $\mcsolve$, given $\mcchc$, synthesizes a partial
  solution $\sigma \in \interps{ \mcpreds }$ that interprets \cc{L3}
  as $\true$, and chooses to then find an interpretation of \cc{L4}.
  %
  The pre-constraint of \cc{L4} is $\true \land \cc{res}_{\cc{4}} =
  0$, where $\cc{res}_{\cc{4}}$ is a logical variable that models the
  value of \cc{res} at line \cc{4}.
\end{ex}

@@

% give the general procedure:
$\shara{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R} }$,
constructs a CDD system $\mathcal{S'}$ equivalent to $\mathcal{S}$.
%
\sys then directly solves $\mathcal{S'}$ and from its solution,
constructs a solution of $\mathcal{S}$.
% define an expansion of a CHC system:
For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in \chcs{
  \mathcal{R} }$ and $\mathcal{S}' \in \chcs{ \mathcal{R}' }$, if
there is a homomorphism from $\mathcal{R}'$ to $\mathcal{R}$ that
preserves the relationship between the clauses of $\mathcal{S}'$ in
the clauses of $\mathcal{S}$, then $\mathcal{S}'$ is an expansion of
$\mathcal{S}$ (all definitions in this section will be over fixed
$\mathcal{R}$, $\mathcal{R}'$, $\mathcal{S}$, and $\mathcal{S}'$).
% formal defn:
\begin{defn}
  \label{defn:expansion}
  Let $f : \mathcal{R}' \to \mathcal{R}$ be such that
  % preserves arity:
  \textbf{(1)} for all $R' \in \mathcal{R}'$, $\arityof{f(R')} =
  \arityof{R'}$;
  % preserves clause structure:
  \textbf{(2)} for all $H \in \apps{\mathcal{R}}$, $\mathcal{B}
  \subseteq \apps{ \mathcal{R} }$, and $\varphi \in \formulas$ such
  that $((\mathcal{B}, \varphi), H) \in \mathcal{S}'$, clause
  %
  \[ \left( \left( \bigunion_{ A \in \mathcal{B} } %
        \{ (f(\relof{A}), \argsof{A}) \}, \varphi \right), %
      ( f(\relof{H}), \argsof{H} ) \right)
  \]
  %
  is in $\mathcal{S}$.
  %
  Then $f$ is a \emph{correspondence} from $\mathcal{S}'$ to
  $\mathcal{S}$.
\end{defn}
%
If there is a correspondence from $\mathcal{S}'$ to $\mathcal{S}$,
$\mathcal{S}'$ is an \emph{expansion} of $\mathcal{S}$, denoted
$\mathcal{S} \expandsto \mathcal{S}'$.
%
\begin{defn}
  \label{defn:min-cdd-expansion}
  If $\mathcal{S}'$ is CDD, %
  $\mathcal{S} \expandsto \mathcal{S}'$, and %
  there is no CDD system $\mathcal{S}''$ such that $\mathcal{S}
  \expandsto \mathcal{S}'' \expandsto \mathcal{S}$ and $\mathcal{S}''
  \not= \mathcal{S}'$, then $\mathcal{S}'$ is a \emph{minimal} CDD
  expansion of $\mathcal{S}$.
\end{defn}

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For $\mathcal{R} \in \relpreds$, $\mathcal{S} \in \chcs{
      \mathcal{R} }$.}
  % Output: inductive invariants.
  \Output{A solution to $\mathcal{S}$ or $\none$.}
  % expand procedure:
  \myproc{$\sys[ \mathcal{R} ](\mathcal{S})$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(\mathcal{S}', \eta) \assign %
    \expand{\mathcal{R}}(\mathcal{S})$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S}' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma' \in \interps{ \ctxs{ \mathcal{R} } }$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{\mathcal{R}}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd{ \ctxs{ \mathcal{R} } }$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
$\shara{ \mathcal{R} }$ (\autoref{alg:shara}), given $\mathcal{S} \in
\chcs{\mathcal{R}}$ (\autoref{line:shara-begin}), returns a solution
to $\mathcal{S}$ or the value $\none$ to denote that $\mathcal{S}$ is
unsolvable.
%
$\shara{ \mathcal{R} }$ first runs a procedure $\expand{ \mathcal{R}
}$ on $\mathcal{S}$ to obtain a minimal CDD expansion $\mathcal{S}'
\in \chcs{ \ctxs{ \mathcal{R} } }$ (\autoref{defn:min-cdd-expansion})
of $\mathcal{S}$ (where $\ctxs{ \mathcal{R} }$ is a space of
relational predicates defined by $\mathcal{R}$, defined in
\autoref{app:cons-cdd}), and a correspondence $\eta : \ctxs{
  \mathcal{R} } \to \mathcal{R}$ (\autoref{line:shara-expand}).
%
$\shara{ \mathcal{R} }$ then runs a procedure $\solvecdd{ \ctxs{
    \mathcal{R} } }$ on $\mathcal{S}'$, which either returns the value
$\none$ to denote that $\mathcal{S}'$ has no solution or a solution
$\sigma'$ (\autoref{line:shara-case}).
%
If $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns that $\mathcal{S}'$ has
no solution, then \sys returns that $\mathcal{S}$ has no solution
(\autoref{line:shara-ret-none}).
%
Otherwise, if $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns a solution
$\sigma'$, then $\sys$ returns, as a solution to $\mathcal{S}$,
$\collapse{ \eta }{ \sigma' }$, defined below
(\autoref{line:shara-ret-collapse}).

% define collapse:
For $\sigma \in \interps{ \ctxs{ \mathcal{R} } }$ and %
correspondence $\eta$ from $\mathcal{S}'$ to $\mathcal{S}$,
$\collapse{\eta}{\sigma} \in \interps{ \mathcal{R} }$ maps each $R \in
\mathcal{R}$ to
%
$\bigland_{ \substack{R' \in \mathcal{R}' \\ \eta(R') = R} }
\sigma(R')$.
%
$\expand{ \mathcal{R} }$ is given in \autoref{app:cons-cdd}.

% shara correctness
\begin{thm}
  \label{thm:corr}
  $\mathcal{S}$ is solvable if and only if $\shara{ \mathcal{R}
  }(\mathcal{S}) \in \interps{ \mathcal{R} }$.
\end{thm}

>>>>

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd(\cc{S})$ %
    \label{line:solve-begin}}{ %
    $\sigma \assign \emptyset$ \\
    $\cc{Preds} \assign \topSort(\predof{S})$ \\
    \For {$\cc{P} \in \cc{Preds}$}
    {$\cc{interpolant} \assign \solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$\\
     \Switch{\cc{interpolant}}{
       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P} \to \cc{I}$]}
     }
    }
    \Return{$\sigma$}
    }
  %
  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
    solution to $\cc{S}$ or the value $\none$ to denote that
    $\cc{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}

% introduce solveCDD
the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
%
$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
uninterpreted predicates in $\cc{S}$ based on their dependency relations.
%
In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
 $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
 encoded pre and post formulas with the current solution.
 %
 If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
 returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
 %
 Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
 interpolation result $\cc{I}$ as part of the soltuion.
 %
 The encoding procedure $\prectr$ and $\postctr$ are defined below.
 %
 $\solvecdd$ returns the soltuion $\sigma$.

\QZ{rework example}
%
\begin{ex}
  $\solvecdd{ \mcpreds }$, given $\mcchc$, may generate
  interpretations of $\mcpreds$ in any topological ordering of the
  dependency relation of $\mcchc$, depending on the relational
  predicates that it chooses at \autoref{line:case-deps}.
  %
  One such ordering is \cc{L3}, \cc{L4}, \cc{dbl}, \cc{L9}, \cc{main},
  $\mathsf{Err}$.
\end{ex}

\subsubsection{Constructing constrains of uninterpreted 
predicate}
%
For an uninterpreted predicate $\cc{P} \in \predof{S}$ and $\cc{S}$ is an CDD system  
,the formula $\ctrof{Q}$ is a compact representation of constrains of $\cc{P}$, defined as follows.
% introduce the isused predicates:
For each $\cc{Q} \in \predof{S}$, there is a corresponding boolean variable
$b_Q$.
% for each relational predicate, define constraint:
The constraint of $P$, denoted $\ctrof{P}$, constrains
that if $P$ occurs in a derivation of $S$, then there is
some clauses $C \in S$ in which $P$ is the head of $\cc{C}$
\textbf{(1)} each $Q \in \bodyof{C}$ is used and 
\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
%
I.e., $\ctrof{P}$ is
\[\lnot b_Q \land 
\left(
\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
\left( \consof{C_i} \land %
\bigland_{ \cc{Q} \in \bodyof{C_i}} %
b_Q \right) \right)
\]
%
The counterexample characterization of $\mathcal{S'}$ constrains that
\textbf{(1)} the query predicate of $\mathcal{S'}$ must be used in a
derivation and %
\textbf{(2)} for $R \in \mathcal{R}$, if $R$ is used, then its
constraint must hold.
%
I.e., $\vc{ \mathcal{R} }(\mathcal{S'})$ is $\used{
  \queryof{\mathcal{S'} } } \land %
\bigland_{R \in \mathcal{R} } \left( \used{R} \implies \ctrof{R}
\right)$.

\QZ{rework example}
\begin{ex}
  % 
  $\vc{ \mcpreds }(\mcchc)$ characterizes all runs of \cc{multiCall}
  that result in an error.
  %
  Although the relational predicates \cc{L4} and \cc{dbl} occur in the
  bodies of multiple clauses, they are each modeled with only a single
  instance of logical variables.
  % 
  Such a collection is sufficient because each run of \cc{multiCall}
  reaches \cc{L4} and \cc{dbl} at most once.
  %
  Correspondingly, each derivation of $\mcchc$ contains at most one
  occurrence of relational predicates \cc{L4} and \cc{dbl}.
\end{ex}

% pre-constraint:
\subsubsection{Constructing the pre-constraint of a relational predicate}
\label{sec:cons-pre}
%
For a CDD system $\cc{S}$, given a uninterpreted predicate $\cc{P} \in \predof{\cc{S}}$
, because $\solvecdd$ solve each uninterpreted predicate in a toplogical order, 
it has a partial solution $\sigma : \cc{Q} \to \formula$ for all uninterpreted predicates in
$\depsof{P}$.
%
the pre-constraint $\prectr(\cc{P},\sigma)$ is a formula that the conjunction of the constrains
of $\cc{P}$ and solutions of all dependencies of $P$ with their boolean varialbe.
%
In particular, $\prectr(\cc{P},\sigma)$ is:
\[\ctrof{P} \land
\left(
\bigland_{\cc{Q} \in \depsof{P}}
\left(b_Q \land \sigma[ \cc{Q} ]  \right)
\right)
\]
%
\QZ{rework example}
\begin{ex}
  \label{ex:pre-ctr}
  Assume that $\mcsolve$, given $\mcchc$, synthesizes a partial
  solution $\sigma \in \interps{ \mcpreds }$ that interprets \cc{L3}
  as $\true$, and chooses to then find an interpretation of \cc{L4}.
  %
  The pre-constraint of \cc{L4} is $\true \land \cc{res}_{\cc{4}} =
  0$, where $\cc{res}_{\cc{4}}$ is a logical variable that models the
  value of \cc{res} at line \cc{4}.
\end{ex}

@@

% give the general procedure:
$\shara$, given an recursion-free CHC system $S$,
constructs a CDD system $S'$ equivalent to $S$.
%
\sys then directly solves $S'$ and from its solution,
constructs a solution of $S$.
% define an expansion of a CHC system:
For two given recursion-free CHC system $S$ and $S'$, if
there is a homomorphism from $\predof{S'}$ to $\predof{S}$ that
preserves the relationship between the clauses of $S'$ in
the clauses of $S$, then $S'$ is an expansion of
$S$ (all definitions in this section will be over fixed
$S$, and $S'$).
% formal defn:
\begin{defn}
  \label{defn:expansion}
  Let $f : \predof{S'} \to \predof{S}$ be such that
  % preserves arity:
  \textbf{(1)} for all $P' \in \predof{S'}$, $P'$ has the
  same vector of parameters as $f(P')$;
  % preserves clause structure:
  \textbf{(2)} for all clause $C' \in S'$, the new clause $C$
  by substituting all predicates $P'$ by $f(P')$ is in $S$.
  %
  Then $f$ is a \emph{correspondence} from $S'$ to $S$.
\end{defn}
%
If there is a correspondence from $S'$ to $S$,
$S'$ is an \emph{expansion} of $S$, denoted
$S \expandsto S'$.
%

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For a recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A solution to $S$ or $\none$.}
  % expand procedure:
  \myproc{$\sys(S)$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(S', \eta) \assign %
    \expand(S)$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd(S' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma'$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
$\shara$ (\autoref{alg:shara}), given an recursion-free CHC system $S$ (\autoref{line:shara-begin}), 
returns a solution
to $S$ or the value $\none$ to denote that $S$ is
unsolvable.
%
$\shara$ first runs a procedure $\expand{}$ on $S$ to obtain a CDD expansion $S'$ (\autoref{defn:min-cdd-expansion})
of $S$.
%
$\shara$ then runs a procedure $\solvecdd$ on $S'$, which either returns the value
$\none$ to denote that $S'$ has no solution or a solution
$\sigma'$ (\autoref{line:shara-case}).
%
If $\solvecdd$ returns that $S'$ has
no solution, then \sys returns that $S$ has no solution
(\autoref{line:shara-ret-none}).
%
Otherwise, if $\solvecdd$ returns a solution
$\sigma'$, then $\sys$ returns, as a solution to $S$,
$\collapse{ \eta }{ \sigma' }$, defined below
(\autoref{line:shara-ret-collapse}).

% define collapse:
For $\sigma'$ and %
correspondence $\eta$ from $P' \in \predof{S'}$ to $P \in \predof{S}$,
$\collapse{\eta}{\sigma'}$ maps solution $\sigma'$ to
$\sigma$[$P \to \bigland_{ \eta(P') = P}
\sigma'(P')$].
%
$\expand{ \mathcal{R} }$ is given in \autoref{app:cons-cdd}.

% shara correctness
\begin{thm}
  \label{thm:corr}
  $S$ is solvable if and only if $\shara$ returns solution $\sigma$.
\end{thm}

### 2018-04-06 paper/background.tex
editing approach, need to rework overview, and move some appendix into main paper

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause contains a body and a head.
%
A body consisting of a conjunctive set of uninterpreted relational predicates applied to varialbes and a formula
that over variables.
%
A head can either be an uninterpreted relational predicates applied to variables or a formula 
relational predicate.
%
A Clause where the head is a formula is called a query.
\begin{defn}
  A Constrained Horn Clause is definied as follows: \\
  $\cc{chc} \Coloneqq \cc{head} \gets \cc{body}$ \\
  $\cc{head} \Coloneqq \varphi \mid \cc{pred}$  \\
  $\cc{body} \Coloneqq \varphi \wedge \cc{preds}$ \\
  $\cc{preds} \Coloneqq \cc{True} \mid \cc{pred} \wedge \cc{preds}$ \\
  $\cc{pred} \Coloneqq$ an uninterpreted predicate applied to variables \\
  $\varphi \Coloneqq$ a formula that all variables are vocabulary \\
\end{defn}
%
% define a body function, head function, and recursion-free
For a given constrained horn clause $C$, $\bodyof{C}$ denotes
the vector of uninterpreted predicates in the body of $C$.
%
It is possible for a uninterpreted predicate shows multiple times in the
body of $C$.
%
For a given constrained horn clause $C$ that is not a query,
$\headof{c}$ denotes the uninterpreted predicates of the head of $C$.
%

@@

\subsubsection{Solutions}
\label{sec:chc-solns}
%
A solution to a CHC system $\mathcal{S}$ is an interpretation of each
relational predicate $R$ of arity $n$ as a formula over $n$ free
variables such that for each clause $\mathcal{C} \in \mathcal{S}$, the
conjunction of interpretations of all relational predicates in the
body of $\mathcal{C}$ and the constraint of $\mathcal{C}$ entail the
interpretation of the head of $\mathcal{C}$.
%
Let a map from each $R \in \relsof{ \mathcal{R} }$ to a
$\mathcal{T}$-formula over an ordered vector of $\arityof{ \mathcal{R}
}(R)$ free variables be an \emph{interpretation} of $\mathcal{R}$;
%
let the space of interpretations of $\mathcal{R}$ be denoted
$\interps{ \mathcal{R} }$.
%
\begin{defn}
  \label{defn:chc-soln}
  For $\mathcal{B} \in \bodies{ \mathcal{R} }$ and $H \in \apps{
    \mathcal{R} }$, %
  let $\sigma \in \interps{ \mathcal{R} }$ be such that for each $R
  \in \mathcal{R}$, $\arityof{\mathcal{R}}(R) = \degreeof{\sigma(R)}$
  and %
  %
  \[ \elts{ \sigma(\relof{ A })[ \argsof{ A } ] }_{ A \in \appsof{
      \mathcal{B} } }, %
  \ctrof{ \mathcal{B} } \entails %
  \sigma( \relof{H} )[ \argsof{ H } ]
  \]
  Then $\sigma$ is a \emph{solution} of $(\mathcal{B}, H)$.
\end{defn}
% define partial solution 
For $\mathcal{S} \in \chcs{ \mathcal{R} }$, if
% solution respects dependence order:
\textbf{(a)} for each $R \in \domain(\sigma)$ (where $\domain(\sigma)$
denotes the domain of $\sigma$) and $R' \in \mathcal{R}$ a dependence
of $R$ in $\mathcal{S}$, it holds that $R' \in \domain(\sigma)$;
%
\textbf{(b)} for each $\mathcal{C} \in \mathcal{S}$ such that
$\relof{ \headof{ \mathcal{C} } } \in \domain(\sigma)$, $\sigma$ is a
solution of $\mathcal{C}$;
%
then $\sigma$ is a \emph{partial solution} of $\mathcal{S}$.
%
If, in addition, $\sigma(\queryof{ \mathcal{S} }) \entails \false$,
then $\sigma$ is a \emph{solution} of $\mathcal{S}$.
%
The problem addressed in this paper is, given a recursion-free CHC
system $\mathcal{S}$, to synthesize a solution for $\mathcal{S}$.

% define grounding of clause body:
Each partial solution $\sigma$ of $\mathcal{S}$ defines a simpler CHC
system in which each relational predicate in the domain of $\sigma$ is
replaced with its interpretation in $\sigma$.
%
For $\mathcal{A} \subseteq \apps{\mathcal{R}}$, $\varphi \in
\formulas$, and $\sigma \in \interps{ \mathcal{R} }$, the clause body
\begin{align*}
  \left( \bigunion_{ \substack{ %
      R \in \mathcal{R} \setminus \domain(\sigma), \\
      Y \in X^{*}, 
      (R, Y) \in \mathcal{A} } } \elts{(R, Y)}, 
  \varphi \land %
  \bigland_{ \substack{ R \in \domain(\sigma) \\ 
      Y \in X^{*},
      (R, Y) \in \mathcal{A} } } \sigma(R)[ Y ] \right)
\end{align*}
%
is the \emph{grounding} of $(\mathcal{A}, \varphi)$ on $\sigma$,
denoted $\sigma(\mathcal{A}, \varphi)$.

% groundings of systems:
For each $\mathcal{S} \in \chcs{ \mathcal{R} }$ and $\sigma \in
\interps{ \mathcal{R} }$ a partial solution of $\mathcal{S}$, the
grounding of $\mathcal{S}$ on $\sigma$ is a CHC system in which each
clause is a grounding of a clause in $\mathcal{S}$ with a head
relational predicate not in the domain of $\sigma$.
%
In particular, let $\mathcal{S}' \subseteq \chc{ \mathcal{R} \setminus
  \domain(\sigma) }$ be such that for each $\mathcal{B} \in \bodies{
  \mathcal{R} }$ and $H \in \apps{ \mathcal{R} }$ with $\relof{H}
\notin \domain(\sigma)$, the clause $(\sigma(\mathcal{B}), H)$ is in
$\mathcal{S}'$.
%
Then $\mathcal{S}'$ is the \emph{grounding} of $\mathcal{S}$ on
$\sigma$, denoted $\sigma(\mathcal{ S })$.

>>>>

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause contains a body and a head.
%
A body consisting of a conjunctive set of uninterpreted relational predicates applied to varialbes and a formula
that over variables.
%
A head can either be an uninterpreted relational predicates applied to variables or a formula 
relational predicate.
%
A Clause where the head is a formula is called a query.
\begin{defn}
  A Constrained Horn Clause is definied as follows: \\
  $\cc{chc} \Coloneqq \cc{head} \gets \cc{body}$ \\
  $\cc{head} \Coloneqq \varphi \mid \cc{pred}$  \\
  $\cc{body} \Coloneqq \varphi \wedge \cc{preds}$ \\
  $\cc{preds} \Coloneqq \cc{True} \mid \cc{pred} \wedge \cc{preds}$ \\
  $\cc{pred} \Coloneqq$ an uninterpreted predicate applied to variables \\
  $\varphi \Coloneqq$ a formula that all variables are vocabulary \\
\end{defn}
%
% define a body function, head function, and recursion-free
For a given constrained horn clause $C$, $\bodyof{C}$ denotes
the vector of uninterpreted predicates in the body of $C$, $\consof{C}$
denotes the formula in the body.
%
It is possible for a uninterpreted predicate shows multiple times in the
body of $C$.
%
For a given constrained horn clause $C$ that is not a query,
$\headof{c}$ denotes the uninterpreted predicates of the head of $C$.
%

@@

%
A solution to a CHC system $S$ is an interpretation of each
uninterpreted predicate $\cc{Q} \in \predof{S}$ that each clause holds
after substitute all predicates by its interpretation.
